Mechanized Reasoning about a Capability Machine

### 2019-10-18 prisc-2020/paper.tex

Next we use the Iris framework to \emph{prove} a Hoare triple for each instruction.
As in \cite{Jensen:2013}, we define a points-to predicate for registers, denoted $r \mapsto_r w$.
Since a capability machine replaces pointers with capabilities, we replace the conventional points-to predicate of separation logic for pointers with a points-to predicate with a permission attached to it, denoted
$a \mapsto_a[p]~w$. 
% \[
%    \hspace{1cm} a \mapsto_a[p]~w.
% \]
This predicate states that address $a$ points to word $w$ with permission $p$.
The permission restricts how the memory may be updated at address $a$.
For instance, if $a \mapsto_a[\textsc{\tiny RX}]~w$, then the $\textsc{\tiny RX}$ (ReadExecute)
permission gives us permission to read from address $a$, to execute w, but it does not permit us to write to address $a$.

% Similar to \cite{Jensen:2013}, the program logic rules for each instruction are of the following form: 
The Hoare triples for basic instructions take the following form
\begin{align*}
\small
	\textit{run time conditions } \wedge &\textit{ decode}(w) = \emph{instr} \Rightarrow  \\
	\{\{\{ \textsf{PC} \mapsto_r ((p,g),b,e,a) &* a \mapsto_a[p] w * ... \}\}\} \\ 
	\textsf{Instr Executable }\\
	\{\{\{\textsf{PC} \mapsto_r ((p,g),b,e,a&+1) * a \mapsto_a[p] w * ... \}\}\}
\end{align*} 
Here, the run time conditions correspond to the dynamic checks done by the capability machine, and \textsf{Instr Executable} is the abstract expression for executing the next instruction in memory.
This form of Hoare triple is similar to the one used in \cite{Jensen:2013} but
unlike \cite{Jensen:2013}, the decoding function is in our case assumed.

>>>>

Next we use the Iris framework to \emph{prove} Hoare triples for each instruction.
As in \cite{Jensen:2013}, we define a points-to predicate for registers, denoted $r \mapsto_r w$.
Since a capability machine replaces pointers with capabilities, we replace the conventional points-to predicate of separation logic for pointers with a points-to predicate with a permission attached to it, denoted
$a \mapsto_a[p]~w$. 
% \[
%    \hspace{1cm} a \mapsto_a[p]~w.
% \]
This predicate states that address $a$ points to word $w$ with permission $p$.
The permission restricts how the memory may be updated at address $a$.
For instance, if $a \mapsto_a[\textsc{\tiny RX}]~w$, then the $\textsc{\tiny RX}$ (ReadExecute)
permission gives us permission to read from address $a$, to execute w, but it does not permit us to write to address $a$.

% Similar to \cite{Jensen:2013}, the program logic rules for each instruction are of the following form: 
The Hoare triples for basic instructions take the following form
\begin{align*}
\small
	\textit{run time conditions } \wedge &\textit{ decode}(w) = \emph{instr} \Rightarrow  \\
	\{\{\{ \textsf{PC} \mapsto_r ((p,g),b,e,a) &* a \mapsto_a[p] w * ... \}\}\} \\ 
	\textsf{Instr Executable }\\
	\{\{\{\textsf{PC} \mapsto_r ((p,g),b,e,a&+1) * a \mapsto_a[p] w * ... \}\}\}
\end{align*} 
Here, the runtime conditions correspond to the dynamic checks done by the capability machine, and \textsf{Instr Executable} is the abstract expression for executing the next instruction in memory.
This form of Hoare triple is similar to the one used in \cite{Jensen:2013} but
unlike \cite{Jensen:2013}, the decoding function is in our case assumed.
